---
title: "How to run a conStruct analysis"
author: "Gideon Bradburd"
date: "2 October 2017"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{run-conStruct}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

<!-- library(rmarkdown) ; render("~/Dropbox/conStruct/code/conStruct/vignettes/format-data.Rmd",html_vignette(toc=TRUE))	-->

## Introduction
This document describes the format of the data used in 
a `conStruct` analysis.

Throughout the document, I'll be referring to the 
example dataset included with the package:

```{r}
library(conStruct)
data(conStruct.data)
```

## conStruct data

There are 3 data objects you need to run a `conStruct` analysis:

1. [allele frequency data]

2. [geographic sampling coordinates]

3. [geographic distance matrix]

In the sections below, I walk through the specific format required for each.

### Allele frequency data

You must specify a matrix of allele frequency data for your samples.
(Make sure the data are of class `matrix`, and that it's not a `data.frame`.)
I assume that the data consist of bi-allelic SNPs.
At each locus, you pick an allele to count across all samples
(it doesn't matter whether it's randomly chosen or 
whether it's always the major or minor allele).
The frequency of the counted allele at a locus in a sample 
is the number of times the counted allele is observed at a locus 
divided by the number of chromosomes genotyped in that sample.
A sample can consist of a single individual or multiple individuals 
lumped together.
So, a successfully genotyped diploid individual heterozygous at a
particular locus would have an allele frequency of 0.5.
If the sample is a population of 13 haploids, of which 12 have the 
counted allele at a given locus, the frequency in that sample at that 
locus would be 12/13.

The matrix of allele frequencies should have one row per sample and 
one column per locus.  Missing data should be denoted with the value `NA`.
An small example allele frequency data matrix is shown below:

| Sample | Locus1 | Locus2 | Locus3 | Locus4 | Locus5 | Locus6 | Locus7 | Locus8 | Locus9 | Locus10 |
|:------|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|
|  Pop1  | 0   | 1    | NA | 0.8 | 0.7 | 0  | 0   | 0.6 | 0   | 1   | 
|  Pop2  | 0   | 1    | 1  | 0.9 | 1   | 1  | 0.1 | 0.6 | 0   | 0.9 |
|  Pop3  | 0.2 | 0.75 | 0  | 1   | 1   | NA | 1   | 1   | 0.1 | 1   |
|  Pop4  | 0.1 | 0.9  | 1  | 1   | 0.8 | 1  | 0.2 | 0.7 | 0.1 | 0.3 |
|  Pop5  | 0   | 1    | 1  | 1   | 1   | 1  | 0.3 | 0.9 | 0.3 | NA  |

An full example allele frequency data matrix is included in the 
`conStruct.data` object included with the package.

```{r}
# load the example data object
data(conStruct.data)

# look at the allele frequency data 
#	for the first 5 populations and 10 loci
conStruct.data$allele.frequencies[1:5,1:10]
```

### Geographic sampling coordinates

You must specify a matrix of geographic sampling coordinates, 
which will be used for plotting the results of the analysis.
This should be a matrix with two columns that give the sample 
x-coordinates (longitude) and y-coordinates (latitude), 
respectively.  The order of rows of the matrix should be the same 
as the order of the rows of the allele frequencies matrix.
If you specify longitude and latitude, they should be in 
decimal degrees.

A full example sampling coordinate data matrix is included 
in the `conStruct.data` object included with the package.

```{r}
# load the example data object
data(conStruct.data)

# look at the geographic sampling coordinates 
#	for the first 5 populations
conStruct.data$coords[1:5,]
```

### Geographic distance matrix 

If you choose to run the spatial model implemented in `conStruct`, 
you must specify a matrix of pairwise geographic distance between 
all samples.  If the coordinates of the samples are real locations 
on Earth (as opposed to simulated coordinates), I recommend 
calculating pairwise great-circle distance between sampling 
coordinates (using, e.g., `fields::rdist.earth`).

The order of the samples in the geographic distance matrix should 
match both that of the geographic coordinates and that of the 
allele frequency data matrix, and all three matrices should have 
the same number of rows.

# Other formats to conStruct

For convenience, I've written a function to convert population 
genetic data in STRUCTURE format to that used in `conStruct`.

## STRUCTURE to conStruct

The program STRUCTURE is one of the most widely used 
methods for model-based clustering in population genetics. 
Many existing programs, including plink (v1.9 and above) 
and PgdSpider, convert data from diverse formats (including 
.vcf files) into STRUCTURE format.  In this section of the 
vignette, I walk through an example of converting a STRUCTURE 
format data file into a `conStruct` format data file.

### STRUCTURE data format
More extensive documentation on STRUCTURE's data format 
can be found in [the STRUCTURE manual](https://web.stanford.edu/group/pritchardlab/structure_software/release_versions/v2.3.4/structure_doc.pdf).
An example STRUCTURE-formatted dataset is shown below:

|        |     | Loc1 | Loc1 | Loc2 | Loc2 | Loc3 | Loc3 | Loc4 | Loc4 |
|:------:|:---:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|:----:|
|  Ind1  |  1  |  1   |  1   |  2   |  2   |   1  |  2   |  -9  |  -9  |
|  Ind2  |  1  |  1   |  2   |  2   |  2   |   2  |  2   |   2  |   2  |
|  Ind3  |  1  |  1   |  1   |  2   |  2   |   2  |  2   |   2  |   2  |
|  Ind4  |  2  |  -9  |  -9  |  1   |  2   |   1  |  1   |   1  |   1  |
|  Ind5  |  2  |  2   |  1   |  2   |  2   |   1  |  1   |   1  |   2  |

	Example STRUCTURE format dataset, with one row per individual 
	and two columns per locus. The first column gives sample names, the 
	second refers to the sample locations, and the last 8 columns give 
	genotype data for four loci. The numbers in the genotype data refer to 
	the allele present at that locus: A1 = `1`, A2 = `2`, missing = `-9`.

To convert a STRUCTURE format file to `conStruct` format, 
you can use the function `structure2conStruct`, included in the 
`conStruct` package.

Below, I give an example of the usage of this function, assuming 
that the file containing the STRUCTURE format data is called 
"myStructureData.str", and that it's on the "desktop" directory 
on the computer.  I also assume that the data are formatted as in 
the table above, with the genotype data starting at the 3rd column 
of the data matrix, and missing data denoted with a value of -9.

```{r,eval=FALSE}
conStruct.data <- structure2conStruct(infile = "~/Desktop/myStructureData.str",
									  start.loci = 3,
				 					  missing.datum = -9,
									  outfile = "~/Desktop/myConStructData")

```

Further documentation for this function is in its help page, 
which you can go to using the command `help(structure2conStruct)`.

If you wish to group multiple individuals together into a single 
sample for analysis you can collapse rows of the `conStruct` format 
data file.  For example, if you have 12 individuals from 4 
locations (3 individuals from each location), and you wish to 
analyze the data treating populations at a sampling location 
as the unit of analysis,  you can do something like the
following:

```{r,eval=FALSE}
pop.data.matrix <- matrix(NA,nrow=4,ncol=ncol(conStruct.data))
for(i in 1:nrow(pop.data.matrix)){
	pop.data.matrix[i,] <- colMeans(
								conStruct.data[
									which(pop.index==i),,
									drop=FALSE
								]
							)
}
```
where `pop.index` is a vector that gives the population of origin 
for each of the individuals sampled.  In the example above, with 
12 individuals sampled from 4 locations (3 from each), 
`pop.index` would be `c(1,1,1,2,2,2,3,3,3,4,4,4)`.


## Microsatellites

This method is designed to run on large datasets consisting of 
bi-allelic SNPs.  If you have a microsatellite dataset and you 
wish to run `conStruct`, the first consideration is whether you 
have sufficient data.  You should have more loci than samples 
in your data matrix (i.e., your data matrix should have more 
columns than rows).

If that's the case, the second consideration is how to format your 
microsat data so that you can run conStruct.  There are two standard 
ways of "SNP-ifying" a microsat dataset.

The first is to lump all microsatellite alleles present at a locus 
into two categories: "major" and "other".  The "major" allele is the 
allele that occurs most frequently at a particular locus; all other 
alleles are put in the "other" bin.  You then can create a dataset 
in which you only report the frequency of the major allele, 
effectively reducing the number of alleles per locus to 2. 
This method has the disadvantage of throwing out data, but 
acknowledges the simplex relationships between alleles at a locus 
(the sum of the frequencies of all alleles at a locus must be 1).

The second approach, introduced by Cavalli-Sforza, is to split out 
each allele at a locus into a separate pseudo-locus consisting of only 
that allele.  That is, if you had 4 alleles present in the genotyped 
sample at a particular locus, at frequencies {0.4,0.3,0.1,0.2}, 
you would split those out into 4 separate columns in your data matrix 
(pseudo-loci), with frequencies in the sampled population of 
{0.4,0.3,0.1,0.2}.  This approach has the advantage of not throwing 
data away, but does not acknowledge the inter-allele dependence 
structure in frequencies, and therefore introduces some 
pseudoreplication into the dataset. This pseudoreplication may make 
you overconfident in your results, as the credible intervals on 
parameter estimates may be artificially narrow.

I would recommend trying both approaches, and comparing the 
estimates of pairwise relatedness you get from each to those 
derived from the raw microsatellite data to see which best 
recovers the patterns of relatedness in the data.  I also recommend 
running `conStruct` on datasets SNP-ified using both approaches, 
and comparing the results.