---
title: "How to run a conStruct analysis"
author: "Gideon Bradburd"
date: "14 December 2017"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{run-conStruct}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

<!-- library(rmarkdown) ; render("~/Dropbox/conStruct/code/conStruct/vignettes/run-conStruct.Rmd",html_vignette(toc=TRUE))	-->

## Introduction
This document describes how to run a `conStruct` analysis.

Throughout the document, I'll be referring to the 
example dataset included with the package:

```{r}
library(conStruct)
data(conStruct.data)
```

The format for the data you need to run a `conStruct` 
analysis is covered in a separate vignette in this 
package. You can view that vignette using the command: 
`vignette(package="conStruct",topic="format-data")`

## Running a conStruct analysis

The function you use to run a `conStruct` analysis 
is called, fittingly, `conStruct`. This vignette 
walks through the use of this function in detail, 
and should be used in concert with the documentation 
for the function, which can be viewed using the command: 
`help(conStruct)`.

### Spatial Model

The default model in the conStruct package is 
the spatial model, which allows relatedness 
within a layer to decay as a function of the distance 
between samples drawing ancestry from that layer.

Below, I show an example of how to run a `conStruct` 
analysis using the spatial model.

```{r,eval=FALSE}
# load the example dataset
data(conStruct.data)

# run a conStruct analysis

#	you have to specify:
#		the number of layers (K)
#		the allele frequency data (freqs)
#		the geographic distance matrix (geoDist)
#		the sampling coordinates (coords)

conStruct(spatial = TRUE, 
	 	  K = 3, 
	  	  freqs = conStruct.data$allele.frequencies,
	  	  geoDist = conStruct.data$geoDist, 
	  	  coords = conStruct.data$coords,
	  	  prefix = "spK3")
```

The function call above runs `conStruct`'s spatial model 
using 3 discrete layers. All output files will be have "spK3" 
prepended to their names. To vary the number of layers 
in the spatial model, you need only change the value of `K`.

### Nonspatial Model

You can also run a nonspatial model using the `conStruct` 
function, in which relatedness within each of the K clusters 
does not decay with distance.  This model is analogous to 
the model implemented in STRUCTURE.

Below, I show an example of how to run a `conStruct` 
analysis using the nonspatial model.

```{r,eval=FALSE}
# load the example dataset
data(conStruct.data)

# run a conStruct analysis

#	you have to specify:
#		the number of layers (K)
#		the allele frequency data (freqs)
#		the sampling coordinates (coords)
#
#	if you're running the nonspatial model, 
#		you do not have to specify 
#		the geographic distance matrix (geoDist)

conStruct(spatial = FALSE, 
	  K = 2, 
	  freqs = conStruct.data$allele.frequencies, 
	  geoDist = NULL, 
	  coords = conStruct.data$coords,
	  prefix = "nspK2")
```

The function call above runs `conStruct`'s nonspatial model 
using 2 discrete layers. All output files will be have "nspK2" 
prepended to their names. As with the spatial model, if you 
want to vary the number of layers, you change the value of `K`.

### Other function options

The `conStruct` function has other arguments that 
have default values, for which you don't have to 
specify any values.  However, you may wish to alter 
these defaults, so we describe them below:

The full function call for the spatial model with 3 layers is:

```{r,eval=FALSE}
conStruct(spatial = TRUE, 
		  K = 3, 
		  freqs = conStruct.data$allele.frequencies, 
		  geoDist = conStruct.data$geoDist, 
		  coords = conStruct.data$coords, 
		  prefix = "spK3", 
		  n.chains = 1, 
		  n.iter = 1000, 
		  make.figs = TRUE, 
		  save.files = TRUE)
```

The other options are `n.chains`, `n.iter`, `make.figs`, `save.files`; 
I describe each of them below:

* `n.chains` - gives the number of independent MCMCs to be run for this model. 
The default is `1`, but you may wish to run multiple independent chains to 
make sure you get consistent results across them.

* `n.iter` - gives the number of iterations per MCMC. The default is `1000`.
If you have more genotyped samples, you generally need more iterations 
to describe the posterior probability surface well. There are no 
hard and fast rules on how many iterations you should run. 
I **strongly recommend** examining model output to assess convergence; 
if you don't see good convergence, you can run the analysis using a 
larger number of iterations.

* `make.figs` - determines whether or not to automatically make figures 
describing the results. The default is `TRUE`. However, if you're running 
lots of independent analyses, or if you're running on a cluster with limited 
disk space, you may wish to set this option to `FALSE` and make the figures 
later on your own.

* `save.files` - determines whether or not to automatically save all output 
files.  The default is `TRUE`.  However, again, there may be circumstances 
in which you don't want to automatically save these files, and instead want 
to capture the results of the analysis, which are the returned value of the 
`conStruct` function call.

## Model diagnosis

As with any statistical model, it is important to assess the 
performance of the inference method. Below, I briefly walk 
through some of the important things to look out for when 
you run a `conStruct` analysis.

### MCMC diagnosis

Although the Hamiltonian Monte Carlo algorithm implemented in STAN 
is quite robust, it's always a good idea to look at the results of 
the analysis to diagnose MCMC performance. If the chain is mixing 
well, the trace plots for the different parameters and the posterior 
probability will resemble a “fuzzy caterpillar,” as in panel (a) 
below. If the trace plots have not plateaued (as in panel (b)), 
it is an indication that the chain has not converged on the 
stationary distribution, and that it should be run longer. 
If the chain appears to be bouncing between two or more modes, 
as in panel (c) below, that may be an indication of a multi-modal 
likelihood surface, with multiple points in parameter space that 
have equal or similar posterior probability given the data. 


```{r,echo=FALSE,fig.width=7,fig.height=2.7}
par(mfrow=c(1,3),mar=c(4,3,1.5,1))
	plot(c(0,rnorm(500,1,0.2)),type='l',
		xlab="",yaxt='n',ylab="")
		mtext(side=2,text="parameter estimate",padj=-1)
		mtext(side=3,text="(a) looks good",padj=-0.1)
	plot(c(0,rnorm(500,c(log(seq(0,1,length.out=500))),0.2)),type='l',
		xlab="",yaxt='n',ylab="")
		mtext(side=1,text="mcmc iterations",padj=2.6)
		mtext(side=3,text="(b) hasn't converged",padj=-0.1)
	plot(c(0,rnorm(150,1,0.2),rnorm(200,3,0.2),rnorm(150,1,0.2)),type='l',
		xlab="",yaxt='n',ylab="")
		mtext(side=3,text="(c) multi-modal",padj=-0.1)
```

### Independent runs

Above, I highlight the importance of evaluating performance of 
individual MCMC runs, but it's also a good idea to run multiple, 
independent analyses and compare results across them.  Ideally, 
multiple independent runs converge on the same stationary distribution, 
with similar parameter estimates and posterior probabilities.  
If different runs give very different results, you can check whether 
there's a mixing problem or a truly multi-modal posterior probability 
surface by comparing the values of the posterior probability across 
runs. If two runs have very different parameter estimates but their 
posterior probability distributions are indistinguishable, that's an 
indication of multi-modality. If multiple runs show different parameter 
estimates, but the posterior probabilities for a subset of the runs that 
show consistent results are higher than those of a different subset that 
gives conflicting results, that indicates that some of the runs are not 
mixing well.

### Missing data

Missing data can affect the sample allelic covariance, and 
therefore the results of a `conStruct` analysis. This is 
especially the case when the distribution of missing data is 
biased - that is, when individuals of particular ancestry are 
more likely to be missing data at a locus. This pattern 
is expected when, for example, allelic dropout occurs in a 
RADseq dataset.

In some empirical datasets with missing data that I used to 
test `conStruct`, I observed a phenomenon of "homogeneous 
minimum layer membership," (HMLM) in which all samples had troublingly 
similar admixture proportions in a particular cluster (see 
membership in the blue layer in the figure below).

\

```{r,echo=FALSE,fig.width=7,fig.height=3}
w <- matrix(rnorm(40,sample(2:10,40,replace=TRUE),1),
			nrow=20,ncol=2)
w <- w/rowSums(w)
w <- cbind(pmax(rnorm(20,0.15,0.005),0),w)
w <- w/rowSums(w)
conStruct::make.structure.plot(w)
```

\

Users are advised to check the results of their analyses carefully 
for this HMLM behavior. If you encounter this issue, try reducing 
the amount of missing data in your dataset, either by dropping 
poorly genotyped samples or poorly genotyped loci (rows and columns 
of the allele frequency data matrix, respectively).